Abb Analysis

### 2017-02-08 abbDataAnalysis_Script.Rmd
rewrite of analysis and viz

  imp.rents <- imputeRents(ltr.df=ltr.data, 
                           str.df=str.data, 
                           mod.spec=ltr.mod.spec,
                           clip.field='suburb')

  str.glob <- merge(str.data, imp.rents$str, by='property.id')

@@

  comp.revs <- compareRevenues(str.imp)

  str.imp <- merge(str.imp, comp.revs, by='property.id')

@@

### Analyze Results

@@

  mrkt.table <- createCompTable(str.imp, 'sub.mrkt')

  mrkt.table$ID <- factor(mrkt.table$ID, 
                          levels=c('city-core', 'city', 'suburban', 'rural', 'beach'))

@@

We plot these results in a 2x2 matrix to show the preference likelihoods under the 4 possible scenarios.

@@

  twotwo.bar.plot <- 
    ggplot(mrkt.table, aes(x=ID, weights=Var, fill=ID)) + 
    geom_bar() +
    facet_grid(~rev.type) +
    scale_fill_manual(values=sm.col) +
    xlab('') +
    ylab('% of properties where Airbnb is more profitable') +
    scale_y_continuous(breaks=c(0,.25,.5,.75,1),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='none')

@@

  rawocc.pplot <- makePrefPlot(str.imp,
                               x.field='pot.occ.rate',
                               y.field='str.ext.pref',
                               group.field='host.type',
                               smooth=TRUE,
                               smooth.span=.75)
  rawocc.pplot <- rawocc.pplot +
    xlab('\nOccupancy Rate') +
    ylab('\n% of Properties where Airbnb is more profitable') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    scale_color_manual(values=sm.col)

@@

  str.revs$sub.mrkt <- factor(str.imp$sub.mrkt, 
                         levels=c('city-core', 'city', 'suburban', 'rural', 'beach'))
  ggplot(str.imp, aes(x=pot.occ.rate, group=sub.mrkt, color=sub.mrkt, fill=sub.mrkt)) +
    geom_density(alpha=.3) +
    facet_wrap(~sub.mrkt) +
    scale_color_manual(values=sm.col) +
    scale_fill_manual(values=sm.col) +
    scale_x_continuous(breaks=seq(0, 1, by=.5),
                       labels=c('0%', '50%', '100%')) +
    ylab('') +
    xlab('Occupancy Rate') +
    theme(legend.position='none',
          axis.line=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks=element_blank(),
          axis.title.y=element_blank())

@@

  str.imp$occ.qtl <- makeWtdQtl(str.imp$occ.rate, 
                                      return.type='rank') 
  str.imp$rate.qtl <- makeWtdQtl(str.imp$med.rate, 
                                  return.type='rank')

@@

  qtlocc.pplot <- makePrefPlot(str.imp,
                               x.field='occ.qtl',
                               y.field='str.act.pref',
                               group.field='sub.mrkt',
                               smooth=TRUE,
                               smooth.span=.75)
  qtlocc.pplot <- qtlocc.pplot +
    xlab('\nQualtile of Occupancy Rate') +
    ylab('\n% of Properties where Airbnb is more profitable') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0', '25th', '50th', '75th', '100th')) +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    scale_color_manual(values=sm.col)

@@

  rate.hm <- makeHeatMap(str.imp,
                x.field='occ.rate',
                y.field='nightly.rate',
                color.field='str.act',
                bins=c(.05, 25),
                svm=F, 
                alpha.count=T,
                add.points=T,
                fill.colors=c(abb.col[1], abb.col[5]))

  rate.hm <- rate.hm +
    xlab('\n Occupancy Rate') +
    ylab('\n Nightly Rate') +
    scale_x_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='bottom')

@@

  rate.hm.svm <- makeHeatMap(str.revs,
                         x.field='occ.rate',
                         y.field='nightly.rate',
                         color.field='str.act',
                         bins=c(.05, 25),
                         svm=T, 
                         alpha.count=F,
                         add.points=T,
                         fill.colors=c(abb.col[1], abb.col[5]))

  rate.hm.svm <- rate.hm.svm +
    xlab('\n Occupancy Rate') +
    ylab('\n Nightly Rate') +
    scale_x_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='bottom')

Again, due to the non-uniformity of distribution of these two axis, we convert to quantiles and re-plot.

  qtl.hm <- makeHeatMap(str.revs,
                         x.field='occ.qtl',
                         y.field='rate.qtl',
                         color.field='str.act',
                         bins=c(5, 5),
                         svm=F, 
                         alpha.count=T,
                         add.points=T,
                         fill.colors=c(abb.col[1], abb.col[5]))

  qtl.hm <- qtl.hm +
    xlab('\n Quantile of Occupancy Rate') +
    ylab('\n Quantile of Nightly Rate') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0', '25th', '50th', '75th', '100th')) +
    scale_y_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0', '25th', '50th', '75th', '100th')) +
    theme(legend.position='bottom')

@@

  qtl.hm.svm <- makeHeatMap(str.revs,
                        x.field='occ.qtl',
                        y.field='rate.qtl',
                        color.field='str.act',
                        bins=c(5, 5),
                        svm=T, 
                        alpha.count=F,
                        add.points=T,
                        fill.colors=c(abb.col[1], abb.col[5]))

  qtl.hm.svm <- qtl.hm.svm +
    xlab('\n Quantile of Occupancy Rate') +
    ylab('\n Quantile of Nightly Rate') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0', '25th', '50th', '75th', '100th')) +
    scale_y_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0', '25th', '50th', '75th', '100th')) +
    theme(legend.position='bottom')

@@

Finally, we calculate the areas of the heat maps and combine into a table.

@@

  svm.rate <- makeSVM(str.revs,
                     x.field='occ.rate',
                     y.field='nightly.rate',
                     z.field='str.act',
                     svm.type='C-svc',
                     svm.kernel='polydot',
                     poly.degree=4,
                     expand.factor=100)

  svm.qtl <- makeSVM(str.revs,
                      x.field='occ.qtl',
                      y.field='rate.qtl',
                      z.field='str.act',
                      svm.type='C-svc',
                      svm.kernel='polydot',
                      poly.degree=4,
                      expand.factor=100)

  market.ratio <- data.frame(type=c('rate', 'qtl'),
                             actual=rep(mean(str.revs$str.act), 2),
                             fitted=c(mean(svm.rate$orig$fitted),
                                      mean(svm.qtl$orig$fitted)),
                             svm=c(mean(svm.rate$pred$pred),
                                   mean(svm.qtl$pred$pred)))
```

### Submarkets

We now take the analytical process just completed on the entire metro and apply it to various submarkets.  We use a custom function that combines all the processes above into a single command.

#### Property Type

We start by comparing apartments to houses.  First we set a new imputation model specification

@@

  # Set model specifications
  ltr.mod.spec <- formula(log(event.price) ~ as.factor(bedbath) + as.factor(suburb) + 
                            as.factor(ltr.month))
  str.mod.spec <- formula(log(nightly.rate) ~ as.factor(bedbath) + as.factor(suburb))

We then recompute the market analysis for each property type.

  # Select data
  str.apt <- which(str.revs$type == 'Apartment')
  ltr.apt <- which(ltr.revs$type == 'Apartment')

  # Estimate House Results  
  house <- fullMarketAnalysis(ltr.df=ltr.revs[-ltr.apt,],
                              str.df=str.revs[-str.apt, ],
                              ltr.mod.spec=ltr.mod.spec,
                              str.mod.spec=str.mod.spec,
                              clip.field='suburb',
                              market.field='none',
                              mrkt.col=sm.col,
                              heat.col=c(abb.col[1], abb.col[5]))

   apt <- fullMarketAnalysis(ltr.df=ltr.revs[ltr.apt,],
                             str.df=str.revs[str.apt, ],
                             ltr.mod.spec=ltr.mod.spec,
                             str.mod.spec=str.mod.spec,
                             clip.field='suburb',
                             market.field='sub.mrkt',
                             mrkt.col=sm.col,
                             heat.col=c(abb.col[1], abb.col[5]))

>>>>

  imp.data <- imputeLtrRents(ltr.df=ltr.data, 
                               str.df=str.data, 
                               mod.spec=ltr.mod.spec,
                               match.factor='suburb')

From the imputed results, we extract the base data (with imputed values) as well as the model summaries.

  # Extract out DFs (only those that imputed)
  str.glob <- merge(str.data, imp.data$imp.rent, by='property.id')

@@

  comp.revs <- compareRevenues(str.glob)

  str.glob <- merge(str.glob, comp.revs, by='property.id')

@@

Realizing that the imputed long term rental values may be more accurate when estimated at smaller scales, we functionalize the above processes and then apply it to short term rentals, broken into submarkets by type, bed/bath and geographic area.  In other words, we separately estimate LTR rents for Apartments with 1bed/1bath in the city core, Apartments with 1bed/1bath in the beach submarket, etc. for all compbinations of type, bed/bath and geographic area. We then combine the results form all estimates back into a single data.frame that contains all of the short-term rentals.

```{r build.tbg.data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 # Create the type, bed/bath and geo sub market variable

 str.data$tbg <- paste0(str.data$product,'.', str.data$sub.mrkt)
 ltr.data$tbg <- paste0(ltr.data$product,'.', ltr.data$sub.mrkt)

 # Make IC estimates at TBG level

 str.data.ic <- abbImputeCompare(str.df=str.data,
                                 ltr.df=ltr.data,
                                 mod.spec=ltr.mod.spec,
                                 match.factor='suburb',
                                 split.field='tbg',
                                 verbose=TRUE)

## Analyze Results

@@

 # Make Basic comparison table
  pref.table <- abbCreateCompTable(ic.df=str.data.ic,
                                   split.field='sub.mrkt')

 # Convert to wide format for output
  pref.table.wide <- dcast(pref.table, ID ~ rev.type, value.var='Var')

 # Limit to the extrapolated pref type
  pref.table <- pref.table[pref.table$rev.type == 'Extrapolated', ]

@@

We then turn this into a simple chart to see the differences

@@

  ggplot(pref.table, aes(x=ID, weights=Var, fill=ID)) + 
    geom_bar() +
    facet_grid(~rev.type) +
    scale_fill_manual(values=sm.col) +
    xlab('') +
    ylab('% of properties where Airbnb is more profitable') +
    scale_y_continuous(breaks=c(0,.25,.5,.75,1),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='none')

@@

  str.data.ic$pref <- str.data.ic$str.ext.pref

  rawocc.pplot <- abbPrefPlot(str.data.ic,
                               x.field='occ.rate',
                               split.field='sub.mrkt',
                               smooth=TRUE,
                               smooth.span=.35 ,
                              spl.col=sm.col)
  rawocc.pplot

@@

  ggplot(str.data.ic, aes(x=occ.rate, group=sub.mrkt, color=sub.mrkt, fill=sub.mrkt)) +
    geom_density(alpha=.3) +
    facet_wrap(~sub.mrkt) +
    scale_color_manual(values=sm.col) +
    scale_fill_manual(values=sm.col) +
    scale_x_continuous(breaks=seq(0, 1, by=.5),
                       labels=c('0%', '50%', '100%')) +
    ylab('') +
    xlab('Occupancy Rate') +
    theme(legend.position='none',
          axis.line=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks=element_blank(),
          axis.title.y=element_blank())

@@

  str.data.ic$occ.qtl <- makeWtdQtl(str.data.ic$occ.rate, 
                                      return.type='rank') 
  str.data.ic$rate.qtl <- makeWtdQtl(str.data.ic$med.rate, 
                                  return.type='rank')

@@

  qtlocc.pplot <- abbPrefPlot(str.data.ic,
                               x.field='occ.qtl',
                               split.field='sub.mrkt',
                               smooth=TRUE,
                               smooth.span=.75)
  qtlocc.pplot <- qtlocc.pplot +
    xlab('\nQualtile of Occupancy Rate') +
    ylab('\n% of Properties where Airbnb is more profitable') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0', '25th', '50th', '75th', '100th')) +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    scale_color_manual(values=sm.col)

  qtlocc.pplot

@@

  rate.hm <- abbHeatMap(str.data.ic,
                x.field='occ.rate',
                y.field='nightly.rate',
                pref.field='str.ext.pref',
                svm=F, 
                alpha.count=T,
                add.points=T,
                fill.colors=c(abb.col[1], abb.col[5]))
  rate.hm

@@

  rate.hm.svm <- abbHeatMap(str.data.ic,
                            x.field='occ.rate',
                            y.field='nightly.rate',
                            pref.field='str.ext.pref',
                            svm=T, 
                            alpha.count=F,
                            add.points=T,
                            fill.colors=c(abb.col[1], abb.col[5]))
  rate.hm.svm

@@

  qtl.hm <- abbHeatMap(str.data.ic,
                x.field='occ.qtl',
                y.field='rate.qtl',
                pref.field='str.ext.pref',
                svm=F, 
                alpha.count=T,
                add.points=T,
                fill.colors=c(abb.col[1], abb.col[5]))
  qtl.hm

@@

@@

  qtl.hm.svm <- abbHeatMap(str.data.ic,
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.ext.pref',
                   svm=T, 
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5]))
  qtl.hm.svm

@@

  svm.rate <- makeSVM(str.data.ic,
                     x.field='occ.rate',
                     y.field='nightly.rate',
                     z.field='str.act.pref',
                     svm.type='C-svc',
                     svm.kernel='polydot',
                     poly.degree=4,
                     expand.factor=100)

  svm.qtl <- makeSVM(str.data.ic,
                      x.field='occ.qtl',
                      y.field='rate.qtl',
                      z.field='str.act.pref',
                      svm.type='C-svc',
                      svm.kernel='polydot',
                      poly.degree=4,
                      expand.factor=100)

  market.ratio <- data.frame(type=c('rate', 'qtl'),
                             actual=rep(mean(str.data.ic$str.act.pref), 2),
                             fitted=c(mean(svm.rate$orig$fitted),
                                      mean(svm.qtl$orig$fitted)),
                             svm=c(mean(svm.rate$pred$pred),
                                   mean(svm.qtl$pred$pred)))
```
